Efficiently stable presentations from error-correcting codes

### 2023-10-04 efficient_stability.tex
Updates from Overleaf
edit

\begin{lemma}\label{lem:lower_bound_on_stability_rate_standard_presentation_Z_2^k}
    Let $\mu_R$ and $\mu_S$ be as in Lemma~\ref{lem:eff-z2}. % be the uniform distribution over the commutation relations in the presentation $\langle S\colon R\rangle$ from \eqref{eq:z2-efficient}, and $\mu_S$ the marginal distribution from $\mu_R$, which is the uniform one on $S=\{x_1,...,x_k\}$. 
    There is a $(\nicefrac{1}{\binom{k}{2}},\mu_R)$-almost homomorphism of $\langle S\colon R\rangle$ which is at least $\nicefrac{1}{16k}$ far from any homomorphism. Namely, the modulus of stability of \eqref{eq:z2-efficient} is $\Omega(k\eps)$.\tnote{Is the `namely'' part immediate? Meaning, why could it not be $\Omega(\sqrt{\eps})$, which for large $\eps$ is better? Maybe you can get an example for a larger range of $\eps$ --- here $\eps$ depends on the size of the presentation. But maybe we can also just leave it at the first sentence.}
\end{lemma}

@@

\tnote{I couldn't follow the last equality above, what allows you to insert $A_1$? An alternate way of going about it is to first show

@@

\big\|\tau(\Id_\mM)-\tau(\Id_\mM B_1^* \Id_\mM B_1 \Id_\mM)\big|
&=\big\|\tau(\Id_\mM A_1^* \Id_\mM A_1 \Id_\mM )-\tau(\Id_\mM B_1^* \Id_\mM B_1 \Id_\mM)\big|\\
&\leq \big| \tau(\Id_\mM (A_1-B_1)^* \Id_\mM A_1\Id_\mM )\big| + \big|\tau(\Id_\mM B_1^*\Id_\mM (A_1-B_1)\Id_\mM )\big|\\
&\leq 2 \big\| \Id_\mM (A_1-B_1)\Id_\mM\big\|_\tau\;,

@@

by Cauchy-Schwwarz. Then, 

@@

  \Vert (A_1-B_1)\Id_\mM\Vert_\tau^2 &= \tau\big(\Id_\mM (A_1-B_1)^*(A_1-B_1)\Id_\mM\big)\\
  &= \tau(\Id_\mM) + \tau(\Id_\mM B_1^*B_1 \Id_\mM) - \tau(\Id_\mM A_1^* B_1 \Id_\mM) - \tau(\Id_\mM B_1^* A_1 \Id_\mM)\\
  &\leq \tau(\Id_\mM) + \tau(\Id_\mM B_1^*\Id_\mM B_1 \Id_\mM) - \tau(\Id_\mM A_1^* \Id_\mM B_1 \Id_\mM) - \tau(\Id_\mM B_1^* \Id_\mM A_1 \Id_\mM) + 2 \big\| \Id_\mM (A_1-B_1)\Id_\mM\big\|_\tau\\
  &\leq 3\Vert \Id_\mM (A_1-B_1)\Id_\mM\Vert_\tau\;,

@@

where the third line uses the previous set of equations to relate the second terms, and $A_1\Id_\mM=A_1$ for the last terms.
This is worse than your bound because of the missing square on the last line, which is due to the first bunch of equations. I think it gives a similar trade-off in the end? 
}

@@

    &\geq \Vert (\Id_\infty-\Id_\mM)B_1A_1 e_v\otimes e_1\Vert_2^2.

@@

\Vert (\Id_\infty -\Id_\mM)B_1A_1\Id_\mM\Vert_\tau^2\leq  \Vert \Id_\mM(\Id_\infty-B_1A_1)\Id_\mM\Vert_\tau^2=\Vert A_1-\Id_\mM B_1\Id_\mM\Vert_\tau^2.

@@

\begin{remark}
As was discussed in the introduction \mnote{Here we should refer to the discussion on the difference between $L^1$ and $L^\infty$ definitions},  it is more common to use the $L^\infty$ analogue of stability. A slight modification of the example given in  the proof of Claim \ref{claim:lower_bound_on_stability_rate_standard_presentation_Z_2^k} provides a lower bound also in that setup. The idea is as follows: In the construction we provided, the pair $(1,2)$ played a special role, since $A_1$ was a CNOT conditioned on the second bit. We can thus define the analogous construction for every pair $(i,j)$ where $1\leq i<j\leq k$. By tensoring all of these constructions along the pairs $(i,j)$, we get $k$ permutations acting on $\binom{[k]}{2}\times \field^{k+1}$, which in turn provide a lower bound in the $L^\infty$ setup. The proof is similar to the above, and since this is a slight digression from our main theme, we omit it from the text. 
\end{remark}

In the next section we will obtain presentations of $\Z_2^k$ that have a much better size/modulus tradeoff.
To formulate the last example we recall the definition of the Pauli matrices
	\begin{equation}\label{eq:def-pauli-1} \sigma^X = \begin{pmatrix} 0 & 1 \\ 1 & 0 \end{pmatrix}\;,\qquad \sigma^Z = \begin{pmatrix} 1 & 0 \\ 0 & -1\end{pmatrix}\;,
	\end{equation}
	and more generally for $a,b\in \Z_2^k$ let 
	\begin{equation}\label{eq:def-pauli-2}
	\sigma^X(a) = \bigotimes_{i=1}^k (\sigma^X)^{a_i}\qquad\text{and}\qquad\sigma^Z(b) = \bigotimes_{i=1}^k (\sigma^Z)^{b_i}\;.
	\end{equation}
These are self-adjoint unitary operators called Pauli observables. Each observable $\sigma^X(a)$ (resp. $\sigma^Z(b)$) corresponds to the \emph{Pauli measurement} $\{ \sigma^X_a \}_{a \in \Z_2^k}$ (resp. $\{ \sigma^Z_b \}_{b \in \Z_2^k}$) where (in a slight abuse of notation)  
	%	We slightly abuse notation and write 
	\[ \sigma^X_a = \Es{\alpha\in\F_2^k} (-1)^{a\cdot \alpha} \sigma^X(\alpha)\qquad\text{and}\qquad\sigma^Z_b = \Es{\beta\in\F_2^k} (-1)^{b\cdot\beta} \sigma^Z(\beta).\]
	It is easy to verify that $\{\sigma^X_a\}_a$ and $\{\sigma^Z_b\}_b$ are projections summing to identity.

>>>>

\begin{lemma}\label{lem:lower_bound_on_stability_rate_standard_presentation_Z_2^k}
    Let $\mu_R$ and $\mu_S$ be as in Lemma~\ref{lem:eff-z2}. % be the uniform distribution over the commutation relations in the presentation $\langle S\colon R\rangle$ from \eqref{eq:z2-efficient}, and $\mu_S$ the marginal distribution from $\mu_R$, which is the uniform one on $S=\{x_1,...,x_k\}$. 
    There is a $(\nicefrac{1}{\binom{k}{2}},\mu_R)$-almost homomorphism of $\langle S\colon R\rangle$ which is at least $\nicefrac{1}{16k}$ far from any homomorphism. Namely, the modulus of stability of \eqref{eq:z2-efficient} is $\Omega(k\eps)$.\tnote{Is the `namely'' part immediate? Meaning, why could it not be $\Omega(\sqrt{\eps})$, which for large $\eps$ is better? Maybe you can get an example for a larger range of $\eps$ --- here $\eps$ depends on the size of the presentation. But maybe we can also just leave it at the first sentence.}\mnote{You are right. I meant that there is no better linear bound... Is there a succinct way of saying that? :)}
\end{lemma}

@@

%\tnote{I couldn't follow the last equality above, what allows you to insert $A_1$? An alternate way of going about it is to first show
%\begin{align*}
%\big\|\tau(\Id_\mM)-\tau(\Id_\mM B_1^* \Id_\mM B_1 \Id_\mM)\big|
%&=\big\|\tau(\Id_\mM A_1^* \Id_\mM A_1 \Id_\mM )-\tau(\Id_\mM B_1^* \Id_\mM B_1 \Id_\mM)\big|\\
%&\leq \big| \tau(\Id_\mM (A_1-B_1)^* \Id_\mM A_1\Id_\mM )\big| + \big|\tau(\Id_\mM B_1^*\Id_\mM (A_1-B_1)\Id_\mM )\big|\\
%&\leq 2 \big\| \Id_\mM (A_1-B_1)\Id_\mM\big\|_\tau\;,
%\end{align*}
%by Cauchy-Schwwarz. Then, 
%\begin{align*}
%  \Vert (A_1-B_1)\Id_\mM\Vert_\tau^2 &= \tau\big(\Id_\mM (A_1-B_1)^*(A_1-B_1)\Id_\mM\big)\\
%  &= \tau(\Id_\mM) + \tau(\Id_\mM B_1^*B_1 \Id_\mM) - \tau(\Id_\mM A_1^* B_1 \Id_\mM) - \tau(\Id_\mM B_1^* A_1 \Id_\mM)\\
%  &\leq \tau(\Id_\mM) + \tau(\Id_\mM B_1^*\Id_\mM B_1 \Id_\mM) - \tau(\Id_\mM A_1^* \Id_\mM B_1 \Id_\mM) - \tau(\Id_\mM B_1^* \Id_\mM A_1 \Id_\mM) + 2 \big\| \Id_\mM (A_1-B_1)\Id_\mM\big\|_\tau\\
%  &\leq 3\Vert \Id_\mM (A_1-B_1)\Id_\mM\Vert_\tau\;,
%\end{align*}
%where the third line uses the previous set of equations to relate the second terms, and $A_1\Id_\mM=A_1$ for the last terms.
%This is worse than your bound because of the missing square on the last line, which is due to the first bunch of equations. %I think it gives a similar trade-off in the end? 
%}
Here the second equality is because by definition, 

@@

    \Vert(\Id_\infty-\Id_\mM)B_1\Id_\mM\Vert_\tau^2=\sum_{\substack{v'\in \field^{k+1} \\ j\neq 1}}\sum_{v\in \field^{k+1}} |(e_{v'}\otimes e_j)^*B_1e_v\otimes e_1|^2=\diamondsuit\;,

@@

but since $A_1$ permutes the basis of $\{e_v\otimes e_1\}$, we have

@@

    \diamondsuit=\sum_{\substack{v'\in \field^{k+1} \\ j\neq 1}}\sum_{v\in \field^{k+1}} |(e_{v'}\otimes e_j)^*B_1A_1e_v\otimes e_1|^2=\Vert (\Id_\infty-\Id_\mM)B_1A_1\Id_\mM\Vert_\tau^2\;.

@@

@@

    &\geq \Vert (\Id_\infty-\Id_\mM)B_1A_1 e_v\otimes e_1\Vert_2^2\;.

@@

\Vert (\Id_\infty -\Id_\mM)B_1A_1\Id_\mM\Vert_\tau^2\leq  \Vert \Id_\mM(\Id_\infty-B_1A_1)\Id_\mM\Vert_\tau^2=\Vert A_1-\Id_\mM B_1\Id_\mM\Vert_\tau^2\;.

@@

\begin{remark}
As was discussed in the introduction \mnote{Here we should refer to the discussion on the difference between $L^1$ and $L^\infty$ definitions},  it is more common to use the $L^\infty$ analogue of stability. A slight modification of the example given in  the proof of Lemma~\ref{lem:lower_bound_on_stability_rate_standard_presentation_Z_2^k} provides a lower bound also in that setup. The idea is as follows: In the construction we provided, the pair $(1,2)$ played a special role, since $A_1$ was a CNOT conditioned on the second bit. We can thus define the analogous construction for every pair $(i,j)$ where $1\leq i<j\leq k$. By tensoring all of these constructions along the pairs $(i,j)$, we get $k$ permutations acting on $\binom{[k]}{2}\times \field^{k+1}$, which in turn provide a lower bound in the $L^\infty$ setup. The proof is similar to the above, and since this is a slight digression from our main theme, we omit it from the text. 
\end{remark}

In the next section we will obtain presentations of $\Z_2^k$ that have a much better size/modulus tradeoff than the one given in~\eqref{eq:z2-efficient}. In the meantime we give one last example. To formulate it we recall the definition of the Pauli matrices
	\begin{equation}\label{eq:def-pauli-1} \sigma^X = \begin{pmatrix} 0 & 1 \\ 1 & 0 \end{pmatrix}\;,\qquad \sigma^Z = \begin{pmatrix} 1 & 0 \\ 0 & -1\end{pmatrix}\;,
	\end{equation}
	and more generally for $a,b\in \Z_2^k$ let 
	\begin{equation}\label{eq:def-pauli-2}
	\sigma^X(a) = \bigotimes_{i=1}^k (\sigma^X)^{a_i}\qquad\text{and}\qquad\sigma^Z(b) = \bigotimes_{i=1}^k (\sigma^Z)^{b_i}\;.
	\end{equation}
These are self-adjoint unitary operators called Pauli observables. Each observable $\sigma^X(a)$ (resp. $\sigma^Z(b)$) corresponds to the \emph{Pauli measurement} $\{ \sigma^X_a \}_{a \in \Z_2^k}$ (resp. $\{ \sigma^Z_b \}_{b \in \Z_2^k}$) where (in a slight abuse of notation)  
	%	We slightly abuse notation and write 
	\[ \sigma^X_a = \Es{\alpha\in\F_2^k} (-1)^{a\cdot \alpha} \sigma^X(\alpha)\qquad\text{and}\qquad\sigma^Z_b = \Es{\beta\in\F_2^k} (-1)^{b\cdot\beta} \sigma^Z(\beta).\]
	It is easy to verify that $\{\sigma^X_a\}_a$ and $\{\sigma^Z_b\}_b$ are projections summing to identity.
